---
title: "Reading Data with Airpred"
author: "Ben Sabath"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
## Settings

Reading data within airpred is controlled by the config file. To generate the config file call:
```{r, eval=FALSE}
gen_config()
```
which will place a formatted config file with the default settings in your current directory.

The fields in the config file you will potentially want to change are `Data_Location`, `Data_Save_Location`, and `finalday`.

### `Data_Location`

This field should contain the path (absolute or relative) to the folder holding the interpolated data
for each monitor. The data should be constructed as one file per year, each file containing a matrix where each column is a monitor and each row represents a day.

### `Data_Save_Location`

The folder where the processed data files should be saved. THis will contain both the long data files for each variable and the final joined dataset.

### `finalday`

This should contain the last day covered by the dataset in the format YYYYMMDD.

## Running the Code

A short R script to assemnble the data would look like the following:

```{R, eval=FALSE}
library(airpred)
process_data()
join_data()
```

The `process data` step takes the separate matrices for each year of a variable and combines them into a single RDS file for each variable. 

The `join_data` step takes each of these files and combines them into a single dataset ordered in a way so that the data could be visualized in a tree structure with each year as the highest level, followed by each day, followed by each monitor. 
